Discrete Quantum Theories and Computing

### 2019-04-29 dissertation.tex
dissertation.lyx Acknowledgments, Preface, "4.7 Quantum Probability Measures over Finite Fields", and "Chapter 6 Conclusion and Further Discussion" Fix small things...

\item My friends living in the Tulip Tree Apartments, including Chao-Hong
Chen (陳昭宏), Tsaiyi Wu (吳采奕), Cheng ``Freddie'' Shi (史程), Ossama
Abdel~Gawwad, and Cathryn ``Cathy'' Creger-Chambers, \textbahasai{Sary
Silvhiany}, Shaozhuan Li (李绍颛), Chenwei Zhang (张晨薇), Shu-Nan ``Nancy''
Chang (張舒涵), and Ching-Chi ``Victoria'' Chang (張靖琪), \ldots , etc.

\item My AT\&T family plan members, especially Meng-Wei Chen (陳孟瑋),
Yao-yu Chih (池耀宇), and Xinyi ``Diana'' Gong (龚欣怡), \ldots , etc.

\item Other friends, including Shu-Ling Wang (王舒齡), Wan-Ling ``Wynnie''
Chang (張婉鈴), Yu-Jung Lin (林宇容), Winnie Lou (何慧芳), Yi-Rong Yang (楊宜蓉),
Yi-Chu Chang (張逸竹), Jim \textgerman[variant=german,spelling=new,babelshorthands=true]{Zimmerly},
Davis Chen (陳鵬舟), Zihang Shao (邵子航), Prof.~Ying Ding (丁颖), Hung-Chun
Chao (趙竑鈞), Guang Zuo (左光), Prof.~Jung-Chao Ban (班榮超), Prof.~Jyh-Chyi
Gong (龔治齊), 江靜宜, and the office of International Services\ldots ,
etc.

@@

\begin{preface*}
\begin{itemize}
\item The latest PDF version of this thesis is in GitHub: \url{https://git.io/fxbuG}~.
\item Its source code is typeset using \LyX{} \cite{LyX} in \url{https://git.io/fxbuE}
and the exported \LaTeX{} file is in \url{https://git.io/fhZBW}~.
\item Any comments can be left in the Issue part of the GitHub repository:
\url{https://git.io/fxbug}~.
\end{itemize}
Following explains how to compile the PDF from the \LaTeX{} source:
\begin{enumerate}
\item \label{enu:Clone-the-git}Clone the git repository of this thesis
\url{https://git.io/fhZP5} into your local directory and the thesis
document class \url{https://github.com/yuttai/iuphd} in a location
where the system can find.
\item In the local directory of this thesis, run \code{xelatex dissertation.tex}~.
\item Run \code{biber dissertation}~.
\item Run \code{xelatex dissertation.tex} several times until all the references
are resolved.
\end{enumerate}
The previous steps only tested in Microsoft Windows with MiK\TeX{}
2.9. Please report me if there is any difficulty to compile it. Besides,
to compile a PDF file from the \LyX{} source, just finish step~\ref{enu:Clone-the-git},
open \code{dissertation.lyx} in \LyX , and click the toolbar button
\includegraphics[width=1em]{buffer-view}.\end{preface*}

@@

After defining a QPMFF~$\mu$, we would like to follow the idea of
Gleason's theorem to see whether we might trace back from $\mu$ to
a pure or mixed state and summarize this correspondence as a discrete
Born rule. When $p=D=3$, there is indeed some correspondence between
QPMFFs and states. However, when $D=3$ and $p=7$, we have numerically
verified that the unique QPMFF $\mu\colon\events_{7^{2}}\rightarrow\left[0,1\right]$
is the equal probable one
\begin{equation}
\mu\left(P\right)=\begin{cases}
0 & \textrm{if }P=\mathbb{0}\,;\\
\frac{1}{3} & \textrm{if }P\textrm{ is a one-dimensional projector;}\\
\frac{2}{3} & \textrm{if }P\textrm{ is a two-dimensional projector;}\\
1 & \textrm{if }P=\mathbb{1}\,.
\end{cases}\label{eq:QPMFF37}
\end{equation}
However, the irreducible pure states in $\ff{7^{2}}^{3}$ is not unique.
Indeed, there are 
\begin{equation}
\frac{7^{3-1}\left(7^{3}-\left(-1\right)^{3}\right)}{7+1}=2107\label{eq:numberOfIrreducibleStates-7-3}
\end{equation}
irreducible pure states in $\ff{7^{2}}^{3}$ as we computed in Eq.~(\ref{eq:numberOfIrreducibleStates}).
Since we don't have enough QPMFF to correspond even pure states, there
is no discrete Born rule in this case. In general, we conjecture that
QPMFF is always unique for $D\ge3$ except $D=p=3$.

@@

Conventional quantum theory is based on the continuum of complex numbers,
but we cannot distinguish two arbitrary complex numbers without unbounded
resources. To explore alternative versions of quantum theory incorporating
our limitation of distinguishability, two types of discrete quantum
theories were described: \emph{quantum theories and computing over
finite fields} and \emph{quantum interval-valued probability measures
(QIVPMs)}. Examining the physical and computational consequences of
such frameworks can yield new insights into not only the subtle properties
of conventional quantum theory but also the power and capacity of
quantum computing.

@@

\begin{figure}[b]
\centering{}$\xymatrix{\textrm{Classical Probability Measure}\ar[d]_{\textrm{blur probability}}\ar[rr]^{\textrm{glue events}} &  & \textrm{Quantum Probability Measure}\ar[d]^{\textrm{blur probability}}\\
\textrm{Classical IVPM}\ar[rr]_{\textrm{glue events}} &  & \textrm{QIVPM}
}
$\caption{\label{fig:commutative-diagrams}QIVPMs inherit from both quantum
probability measures and classical IVPMs.}
\end{figure}
As a natural extension of both conventional quantum probability measures
and classical interval-valued probability measures (IVPMs) illustrated
in Fig.~\ref{fig:commutative-diagrams}, QIVPMs inherit definitions
and properties from the both frameworks. While the expectation values
of QIVPMs can be pulled back to the classical ones and consistent
with those of quantum probability measures in the infinitely precise
limit (Sec.~\ref{sec:Interval-Uncertainty}), foundational concepts
in quantum mechanics, such as the Kochen-Specker and Gleason theorems,
extended to QIVPMs in subtle ways. By carefully specifying experimental
uncertainties, we established rigorous bounds on the validity of the
Kochen-Specker theorem (Sec.~\ref{sec:Kochen-Specker}). While there
is a QIVPM not consistent with Gleason's unique state~$\rho$ on
all projectors, we constructed a class of QIVPMs for which the original
Gleason theorem could be recovered asymptotically (Sec.~\ref{sec:Gleason}).

@@

Therefore, to study a QIVPM~$\bar{\mu}$, we can just study a pair
of functions $\frameL\colon\CP{2}\rightarrow\left[0,1\right]$ and
$\frameR\colon\CP{2}\rightarrow\left[0,1\right]$ defined by $[\frameL\left(\ket{\psi}\right),\frameR\left(\ket{\psi}\right)]=\bar{\mu}\left(\proj{\psi}\right)$

>>>>

\item My friends living in the Tulip Tree Apartments, including Chao-Hong
Chen (陳昭宏), Tsaiyi Wu (吳采奕), Cheng ``Freddie'' Shi (史程), Ossama
Abdel~Gawwad, and Cathryn ``Cathy'' Creger-Chambers, \textbahasai{Sary
Silvhiany}, Shaozhuan Li (李绍颛), Chenwei Zhang (张晨薇), Shu-Han ``Nancy''
Chang (張舒涵), and Ching-Chi ``Victoria'' Chang (張靖琪), \ldots , etc.

\item My AT\&T family plan members, especially Meng-Wei Chen (陳孟瑋),
Yao-Yu Chih (池耀宇), and Xinyi ``Diana'' Gong (龚欣怡), \ldots , etc.

\item Other friends, including Shu-Ling Wang (王舒齡), Wanling ``Wynnie''
Chang (張婉鈴), Yu-Jung Lin (林宇容), Winnie Lou (何慧芳), Yi-Rong Yang (楊宜蓉),
Yi-Chu Chang (張逸竹), Jim \textgerman[variant=german,spelling=new,babelshorthands=true]{Zimmerly},
Peng Chou ``Davis'' Chen (陳鵬舟), Zihang Shao (邵子航), Prof.~Ying Ding
(丁颖), Hung-Chun Chao (趙竑鈞), Guang Zuo (左光), Prof.~Jung-Chao Ban (班榮超),
Prof.~Jyh-Chyi Gong (龔治齊), 江靜宜, and the office of International Services\ldots ,
etc.

@@

\begin{preface*}
\begin{itemize}
\item The latest PDF version of this thesis is in GitHub: \url{https://git.io/fxbuG}~.
\item Its source code is typeset using \LyX{} \cite{LyX} in \url{https://git.io/fxbuE}
and the exported \LaTeX{} file is in \url{https://git.io/fhZBW}~.
\item Any comments can be left in the Issue part of the GitHub repository:
\url{https://git.io/fxbug}~.
\end{itemize}
The following explains how to compile the PDF file from the \LaTeX{}
source:
\begin{enumerate}
\item \label{enu:Clone-the-git}Clone the git repository of this thesis
\url{https://git.io/fhZP5} into your local directory and the thesis
document class \url{https://github.com/yuttai/iuphd} in a location
where the system can find.
\item In the local directory of this thesis, run \code{xelatex dissertation.tex}~.
\item Run \code{biber dissertation}~.
\item Run \code{xelatex dissertation.tex} several times until all the references
are resolved.
\end{enumerate}
The previous steps only tested in Microsoft Windows with MiK\TeX{}
2.9. Please report me if there is any difficulty to compile it. Besides,
to compile a PDF file from the \LyX{} source, just finish step~\ref{enu:Clone-the-git},
open \code{dissertation.lyx} in \LyX , and click the toolbar button
\includegraphics[width=1em]{buffer-view}.\end{preface*}

@@

After defining a QPMFF~$\mu$, we would like to follow the idea of
Gleason's theorem to see whether we might trace back from $\mu$ to
a pure or mixed state and summarize this correspondence as a discrete
Born rule. When $p=D=3$, there is indeed some correspondence between
QPMFFs and states. However, when $D=3$ and $p=7$, we have numerically
verified that the unique QPMFF $\mu\colon\events_{7^{2}}\rightarrow\left[0,1\right]$
is the equal probable one
\begin{equation}
\mu\left(P\right)=\begin{cases}
0 & \textrm{if }P=\mathbb{0}\,;\\
\frac{1}{3} & \textrm{if }P\textrm{ is a one-dimensional projector;}\\
\frac{2}{3} & \textrm{if }P\textrm{ is a two-dimensional projector;}\\
1 & \textrm{if }P=\mathbb{1}\,.
\end{cases}\label{eq:QPMFF37}
\end{equation}
In contrast, the irreducible pure states in $\ff{7^{2}}^{3}$ is not
unique. As we computed in Eq.~(\ref{eq:numberOfIrreducibleStates}),
there are 
\begin{equation}
\frac{7^{3-1}\left(7^{3}-\left(-1\right)^{3}\right)}{7+1}=2107\label{eq:numberOfIrreducibleStates-7-3}
\end{equation}
irreducible pure states in $\ff{7^{2}}^{3}$. Since we don't have
enough QPMFF to correspond even pure states, there is no discrete
Born rule in this case. In general, we conjecture that QPMFF is always
unique for $D\ge3$ except $D=p=3$.

@@

Conventional quantum theory is based on the continuum of complex numbers,
but we cannot distinguish two arbitrary complex numbers without unbounded
resources. To explore alternative versions of quantum theory incorporating
our limitation of distinguishability, two types of discrete quantum
theories were described: \emph{quantum theories and computing over
finite fields} and \emph{quantum interval-valued probability measures
(QIVPMs)}. Examining the physical and computational consequences of
such frameworks could yield new insights into not only the subtle
properties of conventional quantum theory but also the power and capacity
of quantum computing.

@@

\begin{figure}[b]
\centering{}$\xymatrix{\textrm{Classical Probability Measure}\ar[d]_{\textrm{blur probability}}\ar[rr]^{\textrm{glue events}} &  & \textrm{Quantum Probability Measure}\ar[d]^{\textrm{blur probability}}\\
\textrm{Classical IVPM}\ar[rr]_{\textrm{glue events}} &  & \textrm{QIVPM}
}
$\caption{\label{fig:commutative-diagrams}QIVPMs inherit from both quantum
probability measures and classical IVPMs.}
\end{figure}
As a natural extension of both conventional quantum probability measures
and classical interval-valued probability measures (IVPMs) illustrated
in Fig.~\ref{fig:commutative-diagrams}, QIVPMs inherit definitions
and properties from the both frameworks. While the expectation values
with respect to QIVPMs can be pulled back to the classical ones and
consistent with those with respect to quantum probability measures
in the infinitely precise limit (Sec.~\ref{sec:Interval-Uncertainty}),
foundational concepts in quantum mechanics, such as the Kochen-Specker
and Gleason theorems, extended to QIVPMs in subtle ways. By carefully
specifying experimental uncertainties, we established rigorous bounds
on the validity of the Kochen-Specker theorem (Sec.~\ref{sec:Kochen-Specker}).
While there is a QIVPM not consistent with Gleason's unique state~$\rho$
on all projectors, we constructed a class of QIVPMs for which the
original Gleason theorem could be recovered asymptotically (Sec.~\ref{sec:Gleason}).

@@

Therefore, to study a QIVPM~$\bar{\mu}$, we just need to study a
pair of functions $\frameL\colon\CP{2}\rightarrow\left[0,1\right]$
and $\frameR\colon\CP{2}\rightarrow\left[0,1\right]$ defined by $[\frameL\left(\ket{\psi}\right),\frameR\left(\ket{\psi}\right)]=\bar{\mu}\left(\proj{\psi}\right)$
